Fusing Filters with Integer Linear Programming

### 2014-05-04 2014betterfusionforfilters/section/03-SizeInference.tex
fix citation
size inf

Our formulation of size inference is very similar to type inference, a la \CITE.
First, constraints are generated for the bindings, such as equality among two sizes, the conjunction of two constraints, and existentials.
If these constraints can be solved, the equality constraints are used to group the sizes into equivalence classes.
Otherwise if the constraints are unable to be solved, the program cannot be assured to require no runtime size checks and will not be fused by our system.
After the constraints are generated and solved, each combinator is given an iteration size.
Two loops with the same iteration size can be fused together.

@@

The most important constraint is equality on two sizes.
Context holes are used to simplify later definitions, and the notation $\fillhole{c}{d}$ is used to denote replacing all holes in $c$ with $d$.
Constraints may be joined together with $c \wedge d$, which requires both $c$ and $d$ to be satisfiable.
Forall quantifiers may be given an upper bound, and mean the constraint must be satisfiable for every size.
Existential quantifiers simply state that there must exist a size that satisfies the constraint.
\begin{tabbing}
MMMMM   \= MM \= MMMMMMMMM \= \kill
$C$     \> @::=@ \> $true$              \> (trivially true)             \\
%       \> $~|$  \> $\hole$             \> (constraint hole)            \\
        \> $~|$  \> $size = size$       \> (equality constraint)        \\
        \> $~|$  \> $C \wedge C$        \> (conjunction)                \\
        \> $~|$  \> $\forall k_n.\ C$   \> (forall quantification)      \\
        \> $~|$  \> $\exists k_n.\ C$   \> (existential quantification)
\end{tabbing}

>>>>

Our formulation of size inference is very similar to type inference, particularly the exposition of HM(X)~\cite{odersky1999type}.
First, constraints are generated for the bindings, such as equality among two sizes, the conjunction of two constraints, and existentials.
If these constraints can be solved, the equality constraints are used to group the sizes into equivalence classes.
Otherwise if the constraints are unable to be solved, the program cannot be assured to require no runtime size checks and will not be fused by our system.
After the constraints are generated and solved, each combinator is given an iteration size.
Two loops with the same iteration size can be fused together.

@@

The most important constraint is equality on two sizes.
Context holes are used to simplify later definitions, and the notation $\fillhole{c}{d}$ is used to denote replacing all holes in $c$ with $d$.
Constraints may be joined together with $c \wedge d$, which requires both $c$ and $d$ to be satisfiable.
Forall quantifiers mean the constraint must be satisfiable for every size.
Existential quantifiers simply state that there must exist a size that satisfies the constraint.
\begin{tabbing}
MMMMM   \= MM \= MMMMMMMMM \= \kill
$C$     \> @::=@ \> $true$              \> (trivially true)             \\
%       \> $~|$  \> $\hole$             \> (constraint hole)            \\
        \> $~|$  \> $size = size$       \> (equality constraint)        \\
        \> $~|$  \> $C \wedge C$        \> (conjunction)                \\
        \> $~|$  \> $\forall k_n.\ C$   \> (forall quantification)      \\
        \> $~|$  \> $\exists k_n.\ C$   \> (existential quantification)
\end{tabbing}
